---
title: "bounds for beta dist parameters"
author: "Cam Freshwater"
date: "January 31, 2019"
output: html_document
---

Quick simulation exercise to identify reasonable bounds for scale and location parameters in beta distribution for outcome uncertainty. Used to constrain `calcRealCatch()` and reduce errors.

```{r changeMu, echo=TRUE, warning=FALSE}
muSeq <- seq(0.0001, 0.01, length.out = 100)
locationSeq <- rep(NA, length.out = 100)
shapeSeq <- rep(NA, length.out = 100)
realERSeq <- rep(NA, length.out = 100)
refSigma <- 0.07

for (i in seq_along(muSeq)) {
  locationSeq[i] <- muSeq[i]^2 * (((1 - muSeq[i]) / refSigma^2) - (1 / muSeq[i]))
  shapeSeq[i] <- locationSeq[i] * (1 / muSeq[i] - 1)
  realERSeq[i] <- rbeta(1, locationSeq[i], shapeSeq[i], ncp = 0)
}
minValue <- min(which(!is.na(realERSeq)))
locationSeq[minValue-1]
shapeSeq[minValue-1]
muSeq[minValue-1]
refMu <- muSeq[minValue]
refMu #last stable mu
```

Both shape and location parameters need to be non-negative. For the reference sigma value that appears to occur at mu ~= 0.005. Next look at changing sigma.

```{r changeSigma, echo=TRUE, warning=FALSE}
sigmaSeq <- seq(0.005, 0.3, length.out = 100) #unlikely to use sigma below 0.005
locationSeq <- rep(NA, length.out = 100)
shapeSeq <- rep(NA, length.out = 100)
realERSeq <- rep(NA, length.out = 100)

for (i in seq_along(sigmaSeq)) {
  locationSeq[i] <- refMu^2 * (((1 - refMu) / sigmaSeq[i]^2) - (1 / refMu))
  shapeSeq[i] <- locationSeq[i] * (1 / refMu - 1)
  realERSeq[i] <- rbeta(1, locationSeq[i], shapeSeq[i], ncp = 0)
}
maxValue <- max(which(!is.na(realERSeq)))
locationSeq[maxValue+1]
shapeSeq[maxValue+1]
```

Unsurprisingly the minimum value for mu and the maximum for sigma are inversely related. To get around this see if constraining location to be a very small value works.

```{r changeMuTwo, echo=TRUE, warning=FALSE}
refSigma <- 0.07
realERMatrix <- matrix(NA, nrow = 1000, ncol = length(muSeq))
colnames(realERMatrix) <- muSeq
for (i in seq_along(muSeq)) {
  locationSeq[i] <- pmax(0.0001,
                         muSeq[i]^2 * (((1 - muSeq[i]) / refSigma^2) - 
                                         (1 / muSeq[i])))
  shapeSeq[i] <- locationSeq[i] * (1 / muSeq[i] - 1)
  realERMatrix[ , i] <- rbeta(1000, locationSeq[i], shapeSeq[i], ncp = 0)
}
which(is.na(realERMatrix))

##Confirm that it works with larger mu values
bigMuSeq <- seq(0.95, 0.99, length.out = 100)
colnames(realERMatrix) <- bigMuSeq
for (i in seq_along(bigMuSeq)) {
  locationSeq[i] <- pmax(0.0001,
                         bigMuSeq[i]^2 * (((1 - bigMuSeq[i]) / refSigma^2) - 
                                         (1 / bigMuSeq[i])))
  shapeSeq[i] <- locationSeq[i] * (1 / bigMuSeq[i] - 1)
  realERMatrix[ , i] <- rbeta(1000, locationSeq[i], shapeSeq[i], ncp = 0)
}
which(is.na(realERMatrix))
```

Appears to work for both very small and very large target HRs (though an original smaller value of 1e-5 was not stable) confirm with sigma and also check relationship between location parameter and harvest deviations assuming refMu.

```{r changeSigmaTwo, echo=TRUE}
realERMatrix <- matrix(NA, nrow = 1000, ncol = length(sigmaSeq))
colnames(realERMatrix) <- sigmaSeq
for (i in seq_along(sigmaSeq)) {
  locationSeq[i] <- pmax(0.0001,
                         refMu^2 * (((1 - refMu) / sigmaSeq[i]^2) - (1 / refMu)))
  shapeSeq[i] <- locationSeq[i] * (1 / refMu - 1)
  realERMatrix[ , i] <- rbeta(1000, locationSeq[i], shapeSeq[i], ncp = 0)
}

harvDevMat <- (refMu - realERMatrix) / refMu
meanHarvDev <- apply(harvDevMat, 2, mean)
plot(meanHarvDev ~ locationSeq)
```

Confirms that smaller location values give you larger deviations. Therefore coercing location to be small but non-positive shouldn't dramatically underestimate uncertainty. Add 0.0001 constraint to location in calcRealCatch.R function.
